---
title: "Lab 5"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1
Get working directory
```{r}
getwd()
```

# Task 2

## mybin Function
```{r}
mybin=function(iter=100,n=10, p=0.5){
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes in each trial
succ=c()
for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))
#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}
#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))
#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation", xlab="Number of successes")
succ.tab/iter
}
```

## Iterations
```{r}
mybin(n = 10,p = .7,iter = 100)
mybin(n = 10,p = .7,iter = 200)
mybin(n = 10,p = .7,iter = 500)
mybin(n = 10,p = .7,iter = 1000)
obj = mybin(n = 10,p = .7,iter = 10000)
```

## Table
```{r}
obj
dbin<-round(dbinom(x = 0:10,prob = .7,size = 10),4)
names(dbin)=0:10
dbin
```

# Task 3

## Marbles
There is a bag of 20 marbles, 12 white(1) and 8 black(0).

## Sample of marbles

```{r}
sample(c(1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,0),size=5,prob=NULL, replace=F)

sample(c(1,1,1,1,1,1,1,1,1,1,1,1,0,0,0,0,0,0,0,0),size=5,prob=NULL, replace=T)
```

## Hypergeometric Function

```{r}
myhyper=function(iter=100,N=20,r=12,n=5){
# make a matrix to hold the samples
#initially filled with NA's
sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
#Make a vector to hold the number of successes over the trials
succ=c()
for( i in 1:iter){
#Fill each column with a new sample
sam.mat[,i]=sample(rep(c(1,0),c(r,N-r)),n,replace=FALSE)
#Calculate a statistic from the sample (this case it is the sum)
succ[i]=sum(sam.mat[,i])
}
#Make a table of successes
succ.tab=table(factor(succ,levels=0:n))
#Make a barplot of the proportions
barplot(succ.tab/(iter), col=rainbow(n+1), main="HYPERGEOMETRIC simulation", xlab="Number of successes")
succ.tab/iter
}
```

## Iterations

```{r}
myhyper(iter = 100,n = 5,N = 20,r = 12)
myhyper(iter = 200,n = 5,N = 20,r = 12)
myhyper(iter = 500,n = 5,N = 20,r = 12)
myhyper(iter = 1000,n = 5,N = 20,r = 12)
obj = myhyper(iter = 10000,n = 5,N = 20,r = 12)
```

## Table

```{r}
obj
dhyp<-round(dhyper(k = 5,x = 0:5,m = 12,n = 8),4)
names(dhyp) = 0:5
dhyp
```

# Task 4

## Describe mysample

It appears that the mysample function has a sample size of n, an iteration amount, and a specified time. The for loop begins from 1 and increases all the way to the total number of iterations. It then creates a sample and turns that sample into a barplot. Then it appears to create another barplot depending on the amount of time you pass in. Then there will be a total of however many iterations you entered of barplots.


## mysample function

```{r}
mysample=function(n, iter=10,time=0.5){
for( i in 1:iter){
#make a sample
s=sample(1:10,n,replace=TRUE)
# turn the sample into a factor
sf=factor(s,levels=1:10)
#make a barplot
barplot(table(sf)/n,beside=TRUE,col=rainbow(10),
main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
ylim=c(0,0.2)
)

#release the table
Sys.sleep(time)
}
}

```

## Run the function

```{r}
mysample(n = 1000, iter = 1, time = 1)
```

## Description

The function goes through the number of iterations specified, waiting a specified amount of second before outputting the next graph, using a sample size n.

# Task 5

## Choose

```{r}
choose(n = 8,k = 4)
```

## Pois
$$P(Y>4),Y~is~distributed~as~Pois(\lambda=2)$$
P(Y>4)+P(Y<=4)=1, so the lower tail is P(Y>4) = 1-P(Y<=4)
```{r}
1-ppois(lambda = 2,q = 2)
```

## NegBin

```{r}
dnbinom(size = 3,prob = .4,x = 10-3)
```

## Bin

```{r}
pbinom(size = 15,q = 8,prob = .4,lower.tail = T)
```

# Task 6

## Function

```{r}
library(MATH4753Kelly)
MATH4753Kelly::mybin(iter = 1000,n = 5,p = .1)
```

