---
title: "Assessment 1"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

Summarize how I will workout your final grade for the course. Give percentages etc. Give my
grading scale also e.g. What percentage is an A etc.

## Assignments

There will be 4 assignments making up 15% of the grade.

## Laboratories

Roughly 16 will be assigned making up 10% of the grade.

## Projects

2 Projects worth 10% of the grade. Second project is worth twice as much as the first.

## In-Class Quizzes

Worth 10%.

## CANVAS Quizzes

Worth 5%.

## Midterm Exams

Worth 20%.

## Final Exam

Worth 30%.

## Grade Scale

A (90): B (80): C(60): D (50): F (<50)
No curves!!!

# Question 2

## Get data

```{r}
ddt <- read.csv("C:/Users/semyu/OneDrive/Desktop/IDE/RPackages/MATH4753Kelly/MATH4753-Applied Statistical Methods/LABS/Lab 1/DDT.csv")
head(ddt)
```

## Get mile from data frame
```{r}
mile = with(ddt, as.numeric(MILE))
```


## Make coplot

```{r}
coplot(data = ddt, LENGTH~WEIGHT | RIVER * SPECIES, pch = 21, col = mile)
```

## Interpret FCM, LCM, SCM given CCATFISH

What we see in these three graphs is the LENGTH vs WEIGHT of the CCATFISH. The Length of these catfish ALMOST all above 40cm. Their weight has a much wider variance.
```{r}
by(data = ddt$WEIGHT, INDICES = ddt$RIVER, FUN = mean)
```
As we can see the catfish FCM < LCM < SCM. The weight becomes more significant and closer at SCM as opposed to FCM which is more spread out.

## Line A
Line A goes through the data frame and stores the MILE variable from the data frame into variable m.

## Line B
```{r}
length(unique(mile))
```
Line B outputs the number of unique elements in the numeric variable m, or mile in my ".Rmd" file. So 17 is the amount of unique miles where the Corps of Engineers caught the fish.

## 6 plots
The top 6 plots are empty due to the engineers not catching any other fish except CCATFISH in the FCM, LCM, SCM rivers. Since we plotted a conditional plot and those rivers had no other species caught, that is why those 6 are blank.

## Mean
```{r}
fcm.cat <- subset(ddt, RIVER=="FCM" & SPECIES=="CCATFISH")
mean(fcm.cat$DDT)
```

# Question 3
All highway bridges in the US are inspected periodically for deficiency by FHWA. Data from the FHWA inspections are compiled into the NBI. Several of the nearly 100 variables maintained by the NBI are listed below. Classify each variable as quantitative or qualitative.

## Qualitative
c: Toll bridge(Y, N)

e: Condition of deck

g: Route Type

## Quantitative
a: Length of maximum span(feet)

b: Number of lanes

d: Average daily traffic

f: Bypass or detour route

# Question 4
