---
title: "Assessment 1"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

Summarize how I will workout your final grade for the course. Give percentages etc. Give my
grading scale also e.g. What percentage is an A etc.

## Assignments

There will be 4 assignments making up 15% of the grade.

## Laboratories

Roughly 16 will be assigned making up 10% of the grade.

## Projects

2 Projects worth 10% of the grade. Second project is worth twice as much as the first.

## In-Class Quizzes

Worth 10%.

## CANVAS Quizzes

Worth 5%.

## Midterm Exams

Worth 20%.

## Final Exam

Worth 30%.

## Grade Scale

A (90): B (80): C(60): D (50): F (<50)
No curves!!!

# Question 2

## Get data

```{r}
ddt <- read.csv("C:/Users/semyu/OneDrive/Desktop/IDE/RPackages/MATH4753Kelly/MATH4753-Applied Statistical Methods/LABS/Lab 1/DDT.csv")
head(ddt)
```

## Get mile from data frame
```{r}
mile = with(ddt, as.numeric(MILE))
```


## Make coplot

```{r}
coplot(data = ddt, LENGTH~WEIGHT | RIVER * SPECIES, pch = 21, col = mile)
```

## Interpret FCM, LCM, SCM given CCATFISH

What we see in these three graphs is the LENGTH vs WEIGHT of the CCATFISH. The Length of these catfish ALMOST all above 40cm. Their weight has a much wider variance.
```{r}
by(data = ddt$WEIGHT, INDICES = ddt$RIVER, FUN = mean)
```
As we can see the catfish FCM < LCM < SCM. The weight becomes more significant and closer at SCM as opposed to FCM which is more spread out.

## Line A
Line A goes through the data frame and stores the MILE variable from the data frame into variable m.

## Line B
```{r}
length(unique(mile))
```
Line B outputs the number of unique elements in the numeric variable m, or mile in my ".Rmd" file. So 17 is the amount of unique miles where the Corps of Engineers caught the fish.

## 6 plots
The top 6 plots are empty due to the engineers not catching any other fish except CCATFISH in the FCM, LCM, SCM rivers. Since we plotted a conditional plot and those rivers had no other species caught, that is why those 6 are blank.

## Mean
```{r}
fcm.cat <- subset(ddt, RIVER=="FCM" & SPECIES=="CCATFISH")
mean(fcm.cat$DDT)
```

# Question 3
All highway bridges in the US are inspected periodically for deficiency by FHWA. Data from the FHWA inspections are compiled into the NBI. Several of the nearly 100 variables maintained by the NBI are listed below. Classify each variable as quantitative or qualitative.

## Qualitative
c: Toll bridge(Y, N)

e: Condition of deck

g: Route Type

## Quantitative
a: Length of maximum span(feet)

b: Number of lanes

d: Average daily traffic

f: Bypass or detour route

# Question 4

## Simple Random Sample
A simple random sample gives each sample size an equal opportunity of being selected, for a truly random experience.

## Stratified Random Sample
Stratified random sampling is where we separate the population into strata or subgroups, then take a small part of each subgroup to make the sample, to make a more representative sample of the population.

## Cluster Sample
Cluster sampling is where out of the entire population, we take a few clusters or subgroups of the population and analize them in their entirety.

## Systematic Sample
Systematic sampling is where we pick out every kth unit in the population.

# Question 5
NEED MTBE DATA

# Question 6
NEED EQ DATA

# Question 7

## Collection method
The Corps of Engineers used a designed experiment where each subgroup was a different part of the river or a tributary creek.

## Population
The population is the entire Tennessee River and it's Tributary Creeks.

## Qualitative
The River that the fish was caught in, and the species are both qualitative variables in this experiment.

# Question 8

## Type of Graph
Histogram

## Variable measured
Whether the robot has legs, wheels, both, or none. The type of robotic limbs.

## Graph

```{r}
barplot(c(15,8,63,20),xlab = "Types of Robotic Limbs",ylab = "Number of Robots",names.arg = c("None","Both","Legs","Wheels"))
```

From the graph we can see that robots with legs are used the most.

## Relative Frequecy

```{r}
none.rf <- 15/106
both.rf <- 8/106
legs.rf <- 63/106
wheels.rf <- 20/106
print(paste0("None: ", round(none.rf, 4)))
print(paste0("Both: ", round(both.rf,4)))
print(paste0("Legs: ", round(legs.rf,4)))
print(paste0("Wheels: ", round(wheels.rf,4)))
```
## Pareto

```{r}
freq=c(15,8,63,20)
RL=c("None","Both","Legs","Wheels")
l=rep(RL,freq)
pareto<-function(x,mn="Pareto barplot",...){  # x is a vector
x.tab=table(x)
xx.tab=sort(x.tab, decreasing=TRUE,index.return=FALSE)
cumsum(as.vector(xx.tab))->cs
length(x.tab)->lenx
bp<-barplot(xx.tab,ylim=c(0,max(cs)),las=2)
lb<-seq(0,cs[lenx],l=11)
axis(side=4,at=lb,labels=paste(seq(0,100,length=11),"%",sep=""),las=1,line=-1,col="Blue",col.axis="Red")
for(i in 1:(lenx-1)){
segments(bp[i],cs[i],bp[i+1],cs[i+1],col=i,lwd=2)
}
title(main=mn,...)

}
pareto(l)
```


# Question 9

# Question 10

# Question 11

# Question 12

# Question 13

# Question 14

# Question 15
```{r}
library(ggplot2)
g<-ggplot(ddt, aes(x = RIVER,y = LENGTH,fill = SPECIES)) + geom_boxplot() + labs(title = "Samuel Kelly")
g
```

