---
title: "Assingment 4"
author: "Samuel Kelly"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 8/10

# WD

```{r}
getwd()
set.seed(24); y1 = rnorm(30, 10, 5)
 set.seed(34); y2 = rnorm(25, 11, 5)
 t.test(y1,y2, var.equal = TRUE, mu =0, conf.level = 0.90)
 set.seed(24); y1 = rnorm(30, 10, 5)
set.seed(34); y2 = rnorm(25, 11, 5)

t.test(y1,y2, var.equal = TRUE, mu =0, conf.level = 0.95)
set.seed(24); y1 = rnorm(8, 10, 20)
 set.seed(34); y2 = rnorm(10, 11, 21)
 var.test(y1,y2)
 library(s20x)
data("camplake.df")
head(camplake.df)
plot(Length ~ Scale.Radius, data = camplake.df)
y.lm = lm(Length ~ Scale.Radius, data = camplake.df)

normcheck(y.lm, shapiro.wilk = TRUE)
summary(y.lm)
```


# 1. MS 7.118 - pg 364

116 bird species, measured were body mass(g), egg length(mm)

## a.

```{r}
nzbird=read.csv("NZBIRDS.csv")
head(nzbird)
set.seed(50)
sam=nzbird[sample(1:nrow(nzbird),size = 35),]
```

## b.

Body mass

```{r}
mean(sam$Body.Mass)
sd(sam$Body.Mass)
t.test(x = nzbird$Body.Mass,conf.level = .95,mu = 6680.371)$conf.int
```

## c.

We accept our null hypothesis that the mean of our sample, 6680.371g, is in the confidence interval of (3208.217, 13397.434), since 6680.371 is in this interval.

## d.

The true mean of 9113g is in the interval (3208.217, 13397.434). The confidence level we have chosen is alpha=.05, or 95% confidence, and since 95% covers most of the data set, we can be fairly certain that the true mean lies in the interval.

## e.

Egg length

```{r}
mean(sam$Egg.Length,na.rm = T)
sd(sam$Egg.Length,na.rm = T)
t.test(x = nzbird$Egg.Length,conf.level = .95,mu = 55.5303)$conf.int
```

We accept the null hypothesis that 55.5303 is in the interval (53.02102, 68.28667).
61.06 is also in our interval, and since we cover 95% of the data we can be fairly certain that the true mean lies n the interval.

## f.

```{r}
flight_extinct=21/38
flight_alive=7/78
z=qnorm(1-.05/2)
np=38*flight_extinct
np>=4
nq=38*(1-flight_extinct)
nq>=4
np2=78*flight_alive
np2>=4
nq2=78*(1-flight_alive)
nq2>=4
```

Since our approximations are greater than or equal to 4, than we can use CLT.

```{r}
mp=c(-1,1)
(flight_extinct-flight_alive)+mp*z*sqrt((flight_extinct*(1-flight_extinct))/38+(flight_alive*(1-flight_alive))/78)
```

Our 95% ci is (0.2925475, 0.6332285).

## g.

Yes, we 95% confident that 21/38=.5526 lies in the interval while 7/78=.0897 does not, so the proportion of extinct birds is higher.

# 2. MS 7.120 - pg 365

## a.

n1=100, n2=47 are both greater than 30, so we know which 2-sample formula to use.
$$(\bar{y_1}-\bar{y_2})\pm z_{\frac{\alpha}{2}}\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}$$

```{r}
(1312-1352)+mp*qnorm(1-.1/2)*sqrt((422^2)/100+(271^2)/47)
```


## b.

$$(\frac{s_1^2}{s_2^2})\frac{1}{F_{\alpha/2(\nu_1,\nu_2)}}\leq \frac{\sigma_1^2}{\sigma_2^2}\leq (\frac{s_1^2}{s_2^2})F_{\alpha/2(\nu_2,\nu_1)}$$

```{r}
f1=qf(1-.1/2,df1 = 99,df2 = 46)
f2=qf(1-.1/2,df1 = 46,df2 = 99)
l=(422^2)/271^2*(1/f1)
u=(422^2)/271^2*(f2)
ci=c(l,u)
ci
```

We can say that the variances do indeed differ as 1 is not in the interval (1.566267, 3.617750).

# 3. MS 7.128 - pg 367

## a.



## b.



# 4. MS 8.24 - pg 390

## a.

The Null Hypothesis is mu=2, the alternative hypothesis is that mu!=2.

## b.

The t statistic is -1.02, and the p-value is .322

## c.

```{r}
qt(1-.05/2,df = 19)
```


We reject when the t statistic falls in the interval (-infinity,-2.093024)U(2.093024,infinity)

## d.

We accept mu=2 as a possible value. Not only does t lie in the ci, but our p-value is greater than alpha, .322>.05.

## e.

Since our test statistic t lies in the 95% ci, then we can conclude that we are fairly certain our mean lies in the interval.

# 5. MS 8.28 - pg 392

## a.

```{r}
lake=c(9.6,4.5,13.2,4.1,22.6,2.7,14.7,3.5,13.6,19.8,14.3,56.9,25.1,18.4,2.7,4.2,30.2,10.3,17.6,2.4,17.3,38.8,3.0,5.8,7.6)
t.test(x = lake,mu = 15,conf.level = .9)
```

With the t test we can see that our p-value is .8535>.1 so we can accept our null hypothesis, and since the estimate mean is 14.516, we can say that this sample may closely reflect all of Wisconsin.

## b.

$$t=\frac{\bar{y}-\mu_0}{s/\sqrt n}\Rightarrow \bar{y}=t\frac{s}{\sqrt n}+\mu_0$$

```{r}
s=sd(lake)
t=qt(1-.1/2,df = 24)
mu=15
ci=mp*t*(s/sqrt(25))+mu
ci
```

We would reject the null if mu fell into the region (-infinity,10.56413)U(19.43587,infinity).
So the probability that t will lie in that region is:
```{r}
pt((10.56413-14)/(s/sqrt(25)),df = 24) + (1-pt((19.43587-14)/(s/sqrt(25)),df = 24))
```
The probability that t will be in the rejection region is: 12.21687%.

# 6. MS 8.44 - pg 401

```{r}
orch=read.csv("ORCHARD.csv")
head(orch)
library(dplyr)
dim(orch)
filter(orch,CONDITION == "FOG")
filter(orch,CONDITION == "CLEAR"|CONDITION=="CLOUD")
dfog=with(orch,orch[CONDITION == "FOG",])
dfclcloud=with(orch, orch[CONDITION == "CLEAR" | CONDITION == "CLOUD",])
dfog
dfclcloud
fog_rat=subset(orch,CONDITION == "FOG", RATIO)
cc_rat=subset(orch,CONDITION == "CLOUD" | CONDITION == "CLEAR", RATIO)
orch=mutate(orch,Cond=recode(CONDITION,CLOUD="CLCL",CLEAR="CLCL"))
orch
```

```{r}
t.test(RATIO ~ Cond,data = orch,mu=0,var.equal=T)
qt(1-.05/2,df = 8+4-2)
```

Since our t statistic 2.0467 does not fall outside the rejection region of -2.228139 or 2.228139 then we do not reject the null hypothesis that mu0-mu1=0. The p-value also backs this up since .06788>.05.

# 7. MS 8.84 - pg 425 – This refers to 8.39 NOT 8.33!

Variances are equal

## a.

```{r}
gt=read.csv("GASTURBINE.csv")
head(gt)
```

```{r}
gtrad=filter(gt,ï..ENGINE == "Traditional")
gae=filter(gt,ï..ENGINE == "Aeroderiv")
gad=filter(gt,ï..ENGINE == "Advanced")
var.test(x = gtrad$HEATRATE,y = gae$HEATRATE,conf.level = .95,ratio = 1)
qf(c(.05/2,1-.05/2),df1 = 6,df2 = 38)
```

Since F is not our rejection region, and that interval does not contain 1, then we can say that we reject the null hypothesis that sigma1^2/sigma2^2=1. So there is a significant difference between the heat rates between the two different types of engines.

## b.

```{r}
var.test(x = gad$HEATRATE,y = gae$HEATRATE,conf.level = .95,ratio = 1)
qf(c(.05/2,1-.05/2),df1 = 6,df2 = 20)
```

There is an even greater divide between these heat rates, as .05> 1.192e-06, and F is not in our ci. The heat rates of the two engines are significantly different. 

# 8. MS 8.99 - pg 438

## a.



## b.



## c.



## d.



## e.



## f.



# 9. MS 8.104 - pg 439

```{r}
human=c(185.4,146.3,174.4,184.9,240,253.8,238.8,263.5)
robot=c(180.4,248.5,185.5,216.4,269.3,249.6,282,315.9)
t.test(x = human,y = robot,paired = T,conf.level = .95)
qt(c(.05/2,1-.05/2),df = 7)
```

The t test clearly shows that mu1-mu2 does not equal to 0, and this is supported by the p-value .03396<.05, and t lying in the rejection region. Since our estimate is negative, then our mean must be greater for mu2, so the robots clearly took more time than their human counterparts, so the humans were more efficient in their work.

# 10. Plot

```{r}
set.seed(35) # This will give everyone the same sample
sam=round(rnorm(30,mean=20,sd=3),3)
unique(sample(sam,30,replace=TRUE) )

sample(sam,30,replace=FALSE)

myboot<-function(iter=10000,x,fun="mean",alpha=0.05,...){  #Notice where the ... is repeated in the code
n=length(x)   #sample size

y=sample(x,n*iter,replace=TRUE)
rs.mat=matrix(y,nr=n,nc=iter,byrow=TRUE)
xstat=apply(rs.mat,2,fun) # xstat is a vector and will have iter values in it 
ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
# A histogram follows
# The object para will contain the parameters used to make the histogram
para=hist(xstat,freq=FALSE,las=1,
main=paste("Histogram of Bootstrap sample statistics","\n","alpha=",alpha," iter=",iter,sep=""),
...)

#mat will be a matrix that contains the data, this is done so that I can use apply()
mat=matrix(x,nr=length(x),nc=1,byrow=TRUE)

#pte is the point estimate
#This uses whatever fun is
pte=apply(mat,2,fun)
abline(v=pte,lwd=3,col="Black")# Vertical line

t=qt(1-alpha/2,df = n-1)
cit=mp*t*(sd(x)/sqrt(n))+mean(x)

segments(ci[1],0,ci[2],0,lwd=4)      #Make the segment for the ci
text(ci[1],.05,paste("(",round(ci[1],2),sep=""),col="Red",cex=3)
text(ci[2],.05,paste(round(ci[2],2),")",sep=""),col="Red",cex=3)

text(cit[1],.5,paste("(",round(cit[1],2),sep=""),col="Blue",cex=3)
text(cit[2],.5,paste(round(cit[2],2),")",sep=""),col="Blue",cex=3)

# plot the point estimate 1/2 way up the density
text(pte,max(para$density)/2,round(pte,2),cex=3)

return(list(ci=ci,fun=fun,x=x,t=t,cit=cit))# Some output to use if necessary
}

myboot(10000,x=sam,fun=function(x) mean(x),alpha=0.05,xlab="mean(x)",col=rainbow(10))


sam=rnorm(30,mean=20,sd=3)
myboot(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.05,xlab="mean(x)/median(x)",col=rainbow(10))
sam=rchisq(40,df=2)
myboot(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.05,xlab="mean(x)/median(x)",col=rainbow(10))
 

```

