---
title: "Lab 2"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1
Working Directory
```{r}
getwd()
```

## Task 2
Display head of EPAGAS.csv file
```{r}
mpg.df<-read.csv("EPAGAS.csv")
head(mpg.df)
```

## Task 3
  a. Transform the mpg variable to z and verify these results, $$z=0, s=1$$
```{r}
mpg<-mpg.df$MPG
z=(mpg-mean(mpg))/sd(mpg)
layout(matrix(c(1, 2, 3), nrow = 1, ncol = 3))
head(z)
print(paste0("mean = ",round(mean(z), digits = 4)), quote = FALSE)
print(paste0("sd = ",sd(z)),quote=F)
```

  b. Find possible outliers of z.
```{r}
z[abs(z)>3]
mpg[abs(z)>3]
z[abs(z)>=2 & abs(z)<=3]
```
  c. Find actual outliers
```{r}
mpg[abs(z)>3]
```
  
  d. Use lattice package to create dotplot
```{r}
library(lattice)
mpg.col = ifelse(abs(z)>3, "Red",
                 ifelse(abs(z)>=2 & abs(z)<=3, "Blue", "Black"))
dp<-dotplot(mpg,col=mpg.col)
dp
```
  
## Task 4
  
  a. Boxplot of mpg variable
```{r}
bpMPG<- boxplot(mpg, col = "Green", 
                notch = T, main = "Boxplot of MPG", 
                horizontal = T, xlab ="MPG")
```
 
  b. Chebyshev's Rule is $$1-\frac{1}{k^2}$$

  If we are looking the proportion of data that will be in two standard deviations, then we will look at k = 2.
```{r}
cheby <- 1 - (1/2^2)
print(paste0("Proportion of predicted data in two standard deviations is: ",cheby),quote = F)
```
  c. The exact proportion based off of the data within two standard deviations will be at least:
```{r}
length(mpg[abs(z) < 2])/length(mpg)
```
  d. Since 96% of the data is within 2 standard deviations, it agrees with Chebyshev's rule, since at least 75% had to be within 2 standard deviations.
  
  e. The empirical rule requires that are data be unimodal, or a single hump, and be symmetrical around the mode. According to the empirical rule approximately ~95% of data will be within 2 standard deviations.
  
```{r}
plot(density(mpg), main = "MPG Data Set")
abline(v = mean(mpg), col = "Red")
```
  
  f. The data corresponds correctly with the empirical rule as the data set is unimodal, and is symmetrical about the mean.
  
  g. The empirical rule is valid because the data is unimodal, and symmetrical around the mean.
