---
title: "Assignment 3"
author: "Samuel Kelly"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 16/16

# 1. MS 5.54 - pg 215

## a.

rate is 1/mean
mean is in seconds, convert into minutes by 1/60

```{r}
1-pexp(q = 2,rate = 1/(95*1/60))
```

## b.

```{r}
phish <- read.csv("PHISHING.csv")
head(phish)
mu <- mean(phish$INTTIME)
v <-mu^2
mu
sqrt(v)
sd(phish$INTTIME)
```

We can see that the mean of the phishing data set is approximately 5/10ths off from our mean of 95 that we started with, and the sd is exactly the same as the mean when we square it to get the variance, and sqrt it to get the standard deviation. However, performing the sd() function on the data set gets us 91.54 which is a little less than our expected value of standard deviation for an exponential. The data does follow an exponential distribution.

# 2. MS 5.56 - pg 215

## a.

$$\mu=\alpha*\beta$$

```{r}
3*.07
```

$$\sigma^2=\alpha*\beta^2$$

```{r}
3*.07^2
```

## b.

```{r}
c((3*.07)-3*(3*.07^2),(3*.07)+3*(3*.07^2))
```

We should not expect a value of .6 since the interval of .1659,.2541 is 3 sd away from the mean, so we should have at least 8/9 of the data by the Chebyshev rule. The max flood level distribution for the four year period observed should be around .2541 million cubic feet per second.

# 3. MS 5.60 - pg 216

## a.

$$\mu=\alpha*\beta$$

```{r}
2*2
1*4
```

## b.

$$\sigma^2=\alpha*\beta^2$$

```{r}
2*2^2
1*4^2
```

## c.

```{r}
pgamma(1,shape = 2,scale = 2)
pgamma(1,shape = 1,scale = 4)
```

Since formula B gives us a probability of .2212 which is greater than formula A's .0902, then formula B is more likely to give us a reaction in less than 1 minute.

# 4. MS 5.74 - pg 219

## a.

a=a, b=b^(1/a)

```{r}
pweibull(2,2,4^(1/2))
```

## b.

### MEAN

```{r}
4^(1/2)
((2+1)/2)
2*.88623 # gamma value of 1.5 from appendix
```

### sd

```{r}
4^(2/2)
(2+2)/2 #2
(2+1)/2 #1.5
sqrt(4*(1-.88623^2))
```

## c.

```{r}
c(1.77246-2*.926491,1.77246+2*.926491)
```

## d.

It is unlikely that the washing machine will last 6 years, as we need to look at values 5 sd away from the mean to get a value of 6 years.

# 5. MS 5.84 - pg 223

## a.

```{r}
2/(2+9)
(2*9)/(((2+9)^(2))*(2+9+1))
```

## b.

```{r}
1-pbeta(.4,2,9)
```

## c.

```{r}
pbeta(.1,2,9)
```

# 6. MS 5.114 - pg 232

## a.

$$\alpha=2,\beta=16$$

## b.

### Mean

```{r}
(16)^(1/2)
(2+1)/2
4*.88623
```

### variance

```{r}
16^(2/2)
(2+2)/2
(2+1)/2
16*(1-.88623^2)
```

## c.

```{r}
1-pweibull(6,2,16^(1/2))
```


# 7. MS 6.2 - pg 239

## a.

p(x,y)=p(1,1),p(1,2),p(1,3),p(1,4),p(1,5),p(1,6),p(2,1),p(2,2),p(2,3),p(2,4),p(2,5),p(2,6),p(3,1),...p(6,5),p(6,6)=1/36
Each combination has a (1/6)*(1/6)=1/36 chance of popping up.

## b.

p1(x)=p(1)+p(2)+p(3)+p(4)+p(5)+p(6)=1/36+1/36+1/36+1/36+1/36+1/36=1/6

Similarly for p2(y)=1/6

## c.

p1(x|y)
x=1,y=1, (1/36)/(1/6)=1/6
x=1,y=2, 1/36 * 6=1/6
x=1,y=3, 1/36 * 6=1/6
x=1,y=4, 1/36 * 6=1/6
x=1,y=5, 1/36 * 6=1/6
x=1,y=6, 1/36 * 6=1/6
x=2,y=1, 1/36 * 6=1/6
x=2,y=2, 1/36 * 6=1/6
x=2,y=3, 1/36 * 6=1/6
x=2,y=4, 1/36 * 6=1/6
x=2,y=5, 1/36 * 6=1/6
...
x=6,y=5, 1/36 * 6=1/6
x=6,y=6, 1/36 * 6=1/6

p2(y|x)
y=1,x=1, (1/36)/(1/6)=1/6
y=1,x=2, 1/36 * 6=1/6
y=1,x=3, 1/36 * 6=1/6
y=1,x=4, 1/36 * 6=1/6
y=1,x=5, 1/36 * 6=1/6
y=1,x=6, 1/36 * 6=1/6
y=2,x=1, 1/36 * 6=1/6
y=2,x=2, 1/36 * 6=1/6
y=2,x=3, 1/36 * 6=1/6
y=2,x=4, 1/36 * 6=1/6
y=2,x=5, 1/36 * 6=1/6
...
y=6,x=5, 1/36 * 6=1/6
y=6,x=6, 1/36 * 6=1/6

## d.

Since p1(x)=p1(x|y) and p2(y)=p2(y|x) then the variables are independent.

# 8. MS 6.4 - pg 240

## a.

```{r}
gran <- matrix(c(1/7,0,0,2/7,0,0,1/7,2/7,1/7),byrow = F,nrow = 3,ncol = 3)
colnames(gran) <- c("1 Energy","2 Energy","3 energy")
rownames(gran) <- c("1 time", "2 time", "3 time")
gran <- as.table(gran)
gran
```

## b.

```{r}
addmargins(gran)
```

The marginal distribution p1(x) is 1/7~.1429, 2/7~.2857, 4/7~.5714

## c.

The marginal distribution p2(y) is 4/7~.5714, 2/7~.2857, 1/7~.1429

## d.

```{r}
#p2(y|x) is p2(1|1)= 1/7/1/7=1/7*7=1
#p2(1|2)= 0/7/1/7=0/7*7=0
#p2(1|3)= 0/7/1/7=0/7*7=0
#p2(2|1)= 2/7/2/7=2/7*7/2=1
#p2(2|2)= 0/7/2/7=0/7*7/2=0
#p2(2|3)= 0/7/2/7=0/7*7/2=0
#p2(3|1)= 1/7/4/7=1/7*7/4=1/4
#p2(3|2)= 2/7/4/7=2/7*7/4=1/2
#p2(3|3)= 1/7/4/7=1/7*7/4=1/4
```

# 9. MS 6.12 - pg 244

## a.

$$f(x,y)=\frac{e^{\frac{-y}{10}}}{10y},0<y<x<2y$$
$$f(y)=\int_{-\infty}^{\infty}f(x,y)dx$$
$$f(y)=\int_{y}^{2y}\frac{e^{\frac{-y}{10}}}{10y}dx$$
$$f(y)=\frac{e^{\frac{-y}{10}}}{10y}\int_{y}^{2y}dx$$
$$f(y)=\frac{e^{\frac{-y}{10}}}{10y}(2y-y)=\frac{e^{\frac{-y}{10}}}{10y}y=\frac{e^{\frac{-y}{10}}}{10}$$

It appears to be an exponential distribution.
$$f(y)=\frac{e^{\frac{-y}{\beta}}}{\beta}$$

## b.

$$E(Y)=\mu=\beta=10$$

# 10. MS 6.14 - pg 245

## a.

$$\int_{-\infty}^{\infty}\int_{-\infty}^{\infty}f(x,y)dxdy=1$$
$$\int_{0}^{x}\int_{0}^{\infty}ce^{-x^2}dxdy=1$$
$$c\int_{0}^{\infty}\int_{0}^{x}e^{-x^2}dydx=1$$
$$c\int_{0}^{\infty}e^{-x^2}y|_{0}^xdx=1$$
$$c\int_{0}^{\infty}e^{-x^2}(x)dx=1$$
$$c\int_{0}^{\infty}\frac{-1}{2x}xe^{u}du=1,u=-x^2,du/dx=-2x$$
$$-\frac{1}{2}c\int_{0}^{\infty}e^{u}du=1$$
$$-\frac{1}{2}ce^{-x^2}|_0^\infty=1$$
$$-\frac{1}{2}c(e^{-\infty^2}-e^0)=1$$
$$-\frac{1}{2}c(0-1)=1$$
$$-\frac{1}{2}c(-1)=1$$
$$c=2$$

## b.

$$f_1(x)=\int_{-\infty}^\infty f(x,y)dy$$
$$f_1(x)=\int_{0}^x 2e^{-x^2}dy$$
$$f_1(x)= 2e^{-x^2}y|_0^x$$
$$f_1(x)= 2e^{-x^2}(x-0)$$
$$f_1(x)= 2xe^{-x^2}$$
$$\int_{-\infty}^\infty f_1(x)dx=\int_{0}^\infty 2xe^{-x^2}dx=\int_{0}^\infty \frac{-1}{2x}2xe^{u}du,u=-x^2,du/dx=-2x,dx=\frac{-1}{2x}du$$
$$-\int_{0}^\infty e^{u}du$$
$$-e^{-x^2}|_0^\infty=-(e^{-\infty}-e^0)=-(0-1)=--1=1$$

## c.

$$f_2(y|x)=\frac{f(x,y)}{f_1(x)}$$
$$f_1(x)=2xe^{-x^2},0\leq x<\infty$$
$$f(x,y)=2e^{-x^2},0\leq y\leq x,0\leq x<\infty$$
$$f_2(y|x)=\frac{2e^{-x^2}}{2xe^{-x^2}}=\frac{1}{x}$$
$$\text{A uniform distribution is }f(y)=\frac{1}{b-a},a\leq y \leq b.$$
$$\frac{1}{x-0},0\leq y\leq x$$

# 11. MS 6.51 - pg 253

$$\text{If two random variables X and Y are independent then Cov(X,Y)=0}$$
The Cov(X,Y)=E(XY)-E(X)E(Y)=0, so if Cov(X,Y)=0 then we need to show that X and Y dependent. So E(XY)!=E(X)E(Y).
So marginals of X are 4/12,4/12,4/12 and Y is 4/12,4/12,4/12.
P(-1,-1)=P(-1)P(-1), 1/12=4/12*4/12=8/12
P(-1,0)=P(-1)P(0), 2/12=4/12 * 4/12=8/12
Since P(X,Y)!=P(X)P(Y) then cov=0 does not imply that X and Y are independent.

# 12. MS 6.74 - pg 269

## a.

```{r}
(3+1)/2
```

The mean of a uniform distribution is (a+b)/2, so (min+max)/2, so the average is 2 parts per million of uranium.

## b.

```{r}
((3-1)^2)/(12*60)
```

## c. 

By CLT with n large enough, we will see a Normal distribution.

## d.

```{r}
pnorm(2.5,mean = 2,sd = sqrt(((3-1)^2)/(12*60)))-pnorm(1.5,mean = 2,sd = sqrt(((3-1)^2)/(12*60)))
```

## e.

```{r}
pnorm(2.2,mean = 2,sd = sqrt(((3-1)^2)/(12*60)))
```


# 13. MS 6.90 - pg 273

## a.

```{r}
pnorm(1.5,mean = 20*.4,sd = sqrt(20*.4*.6))
```

## b.

```{r}
1-pnorm(10.5,mean = 20*.4,sd = sqrt(20*.4*.6))
```

## c.

```{r}
pbinom(1,size = 20,prob = .4)
1-pbinom(10,size = 20,prob = .4)
```

The normal approximation is pretty close in the second case, but further away in the first case, but I would say the approximation is close enough.

# 14. MS 7.108 - pg 362

## a.

```{r}
t.test(x = c(1.32,0,13.1,.919,.657,3,1.32,4.09,4.45,0),conf.level = .99)$conf.int
```

## b.

```{r}
t.test(x = c(.508,.279,.32,.904,.221,.283,.475,.13,.22,.743),conf.level = .99)$conf.int
```

## c.

The confidence intervals show that the mean value most likely falls in this interval. So the mean lead level of the water is in the interval (-1.15,6.92) and similarly for copper which is between (.15,.66).

## d.

We are 99% confident that our mean lies within these intervals. In the interval there is a 99% certainty that the interval contains the true value that we are looking for.

# 15. MS 7.114 - pg 364

```{r}
y1 = c(782,965,948,1181,1414,1633,1852)
y2 = c(593,672,750,988,1226,1462,1698)
t.test(y1,y2,conf.level = .95,paired = T)$conf.int
```

We are looking at the confidence interval of the difference of 2 mean values where the interval tells us the average difference between the two locations.

# 16. MS 7.116 - pg 364

## a.

```{r}
y1 = c(5.4,2.7,34.2,19.9,2.4,7,6.1,7.7,18.4,27.1,16.9)
y2 = c(24.3,16.5,47.2,12.4,24,21.6,104.3,96.9,105.3,78.7,44.6)
t.test(y1,y2,conf.level = .9,paired = T)$conf.int
```

## b.

We are assuming that the values are paired which is true since they depend on each other.

## c.

The mean difference of diazinon between day and night is definitely different. We have a confidence interval of (-58.89917, -18.91901), which contains the average variability between day and night, and the night values are larger than the day values.
