---
title: "Assignment 3"
author: "Samuel Kelly"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. MS 5.54 - pg 215

## a.

rate is 1/mean
mean is in seconds, convert into minutes by 1/60

```{r}
1-pexp(q = 2,rate = 1/(95*1/60))
```

## b.

```{r}
phish <- read.csv("PHISHING.csv")
head(phish)
mu <- mean(phish$INTTIME)
v <-mu^2
mu
sqrt(v)
sd(phish$INTTIME)
```

We can see that the mean of the phishing data set is approximately 5/10ths off from our mean of 95 that we started with, and the sd is exactly the same as the mean when we square it to get the variance, and sqrt it to get the standard deviation. However, performing the sd() function on the data set gets us 91.54 which is a little less than our expected value of standard deviation for an exponential. The data does follow an exponential distribution.

# 2. MS 5.56 - pg 215

## a.

$$\mu=\alpha*\beta$$

```{r}
3*.07
```

$$\sigma^2=\alpha*\beta^2$$

```{r}
3*.07^2
```

## b.

```{r}
c((3*.07)-3*(3*.07^2),(3*.07)+3*(3*.07^2))
```

We should not expect a value of .6 since the interval of .1659,.2541 is 3 sd away from the mean, so we should have at least 8/9 of the data by the Chebyshev rule. The max flood level distribution for the four year period observed should be around .2541 million cubic feet per second.

# 3. MS 5.60 - pg 216

## a.

$$\mu=\alpha*\beta$$

```{r}
2*2
1*4
```

## b.

$$\sigma^2=\alpha*\beta^2$$

```{r}
2*2^2
1*4^2
```

## c.

```{r}
pgamma(1,shape = 2,scale = 2)
pgamma(1,shape = 1,scale = 4)
```

Since formula B gives us a probability of .2212 which is greater than formula A's .0902, then formula B is more likely to give us a reaction in less than 1 minute.

# 4. MS 5.74 - pg 219



# 5. MS 5.84 - pg 223



# 6. MS 5.114 - pg 232



# 7. MS 6.2 - pg 239



# 8. MS 6.4 - pg 240



# 9. MS 6.12 - pg 244



# 10. MS 6.14 - pg 245



# 11. MS 6.51 - pg 253



# 12. MS 6.74 - pg 269



# 13. MS 6.90 - pg 273



# 14. MS 7.108 - pg 362



# 15. MS 7.114 - pg 364



# 16. MS 7.116 - pg 364


