---
title: "Lab 12"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

```{r}
getwd()
```

# Task 2

## Create Sample

Use x1 from parallel code to form our sample.

```{r}
set.seed(55);
x1=rnorm(30,mean=25,sd=5)
```

## Test mean 22

```{r}
t22=t.test(x = x1,mu = 22)
t22$p.value
t22$conf.int
t22$statistic
qt(1-.05/2,29)
```

Since our default value of alpha=.05 which is greater than our p-value of .00205 then we reject the null hypothesis of mu=22. An alternative way to reject is to see if 22 lies in the interval (23.302,27.2732), which it does not, so we still reject the null hypothesis. The third way to reject is to see if the absolute value of our t statistic is greater than the critical value we calculate, and 3.386 > 2.045, so we reject the null hypothesis.

## Test mean 23

```{r}
t23=t.test(x = x1,mu = 23)
t23$p.value
t23$conf.int
#t23$statistic
#qt(1-.05/2,29)
```

We reject the null as .05>.002543 and 23 is not the interval (23.30198, 27.27320).

## Test mean 24

```{r}
t24=t.test(x = x1,mu = 24)
t24$p.value
t24$conf.int
#t24$statistic
#qt(1-.05/2,29)
```

We accept the null hypothesis since .05 is not greater than .1951074, and 24 is in the interval (23.30198, 27.27320).

## Test mean 25

```{r}
t25=t.test(x = x1,mu = 25)
t25$p.value
t25$conf.int
#t22$statistic
#qt(1-.05/2,29)
```

The greater the p-value the better. Here .7691681 > .05 so we accept the null hypothesis and 25 is in the interval (23.30198, 27.27320).

## Test mean 26

```{r}
t26=t.test(x = x1,mu = 26)
t26$p.value
t26$conf.int
#t26$statistic
#qt(1-.05/2,29)
```

Accept null hypothesis since .468963 > .05 and 26 is in the interval (23.30198, 27.27320).

## Boxplot

```{r}
boxplot(x1, main="Sample x1")
ci=t.test(x1,mu=23)$conf.int
ci
abline(h=c(ci,mean(x1)),col=c("Red","Red","Green"))
```

The green line is our sample mean, while the red lines are our confidence interval, or rejection region.

## tcalc

$$t_{calc}=\frac{\bar{x1}-\mu_0}{\frac{s}{\sqrt{n}}}$$

```{r}
tcalc=(mean(x1)-24)/(sd(x1)/sqrt(30))
tcalc
t24=t.test(x = x1,mu = 24)
t24$statistic
```

Both t values equal each other.

## mypvalue function

```{r}
# Display P-value areas
mypvalue=function(t0,xmax=4,n=20, alpha=0.05){
#calculate alpha/2
va=round(pt(-t0,df=n-1),4)
pv=2*va

# plot the t dist
curve(dt(x,df=n-1),xlim=c(-xmax,xmax),ylab="T Density",xlab=expression(t),
main=substitute(paste("P-value=", pv, " alpha=", alpha)))


# set up points on the polygon to the right
xcurve=seq(t0,xmax,length=1000)
ycurve=dt(xcurve,df=n-1)

# set up points to the left
xlcurve=seq(-t0,-xmax,length=1000)
ylcurve=dt(xcurve,df=n-1)

# Shade in the polygon defined by the line segments
polygon(c(t0,xcurve,xmax),c(0,ycurve,0),col="green")
polygon(c(-t0,xlcurve,-xmax),c(0,ylcurve,0),col="green")

# make quantiles
q=qt(1-alpha/2,n-1)
abline( v=c(q,-q),lwd=2) # plot the cut off t value 
axis(3,c(q,-q),c(expression(abs(t[alpha/2])),expression(-abs(t[alpha/2]))))


# Annotation
text(0.5*(t0+xmax),max(ycurve),substitute(paste(area, "=",va)))
text(-0.5*(t0+xmax),max(ycurve),expression(area))

return(list(q=q,pvalue=pv))
}
```

## Plot

```{r}
mypvalue(tcalc,n = 30,alpha = .05)
```

The rejection region is the values that the function qt generates, so the interval of (-infinity,-2.04523)U(2.04523,infinity).
Our p-value is .1952.
The tcalc that we calculated is 1.326252 which is not in the rejection region.

## Bootstrap function

```{r}
bootpval<-function(x,conf.level=0.95,iter=3000,mu0=0, test="two"){
n=length(x)
y=x-mean(x)+mu0  # transform the data so that it is centered at the NULL
rs.mat<-c()    #rs.mat will become a resample matrix -- now it is an empty vector
xrs.mat<-c()
for(i in 1:iter){ # for loop - the loop will go around iter times
rs.mat<-cbind(rs.mat,sample(y,n,replace=TRUE)) #sampling from y cbind -- column bind -- binds the vectors together by columns
xrs.mat<-cbind(xrs.mat,sample(x,n,replace=TRUE)) #sampling from x cbind -- column bind -- binds the vectors together by columns

}

tstat<-function(z){ # The value of t when the NULL is assumed true (xbar-muo)/z/sqrt(n)
sqrt(n)*(mean(z)-mu0)/sd(z)
}

tcalc=tstat(x) # t for the data collected
ytstat=apply(rs.mat,2,tstat) # tstat of resampled y's, ytstat is a vector and will have iter values in it
xstat=apply(xrs.mat,2,mean)  # mean of resampled x's
alpha=1-conf.level # calculating alpha
ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
pvalue=ifelse(test=="two",length(ytstat[ytstat>abs(tcalc) | ytstat < -abs(tcalc)])/iter,
ifelse(test=="upper",length(ytstat[ytstat>tcalc])/iter,
length(ytstat[ytstat<xstat])/iter))

h=hist(ytstat,plot=FALSE)
mid=h$mid
if(test=="two"){
ncoll=length(mid[mid<= -abs(tcalc)])
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll-ncolr),rep("Green",ncolr))
}
if(test=="upper"){
ncolr=length(mid[mid>=  abs(tcalc)])
col=c(rep("Gray",length(mid)-ncolr),rep("Green",ncolr))
}

if(test=="lower"){
ncoll=length(mid[mid<=  -abs(tcalc)])
col=c(rep("Green",ncoll),rep("Gray",length(mid)-ncoll))
}
hist(ytstat,col=col,freq=FALSE,las=1,main="",xlab=expression(T[stat]))
#segments(ci[1],0,ci[2],0,lwd=2)
pround=round(pvalue,4)
title(substitute(paste(P[value],"=",pround)))
return(list(pvalue=pvalue,tcalc=tcalc,n=n,x=x,test=test,ci=ci))
}
```

## Test mean 22

```{r}
bootpval(x = x1,mu0 = 22,test = "two")
```

## Test mean 23

```{r}
bootpval(x = x1,mu0 = 23,test = "two")
```

## Test mean 24

```{r}
bootpval(x = x1,mu0 = 24,test = "two")
```

## Test mean 25

```{r}
bootpval(x = x1,mu0 = 25,test = "two")
```

## Test mean 26

```{r}
bootpval(x = x1,mu0 = 26,test = "two")
```

The above output compared to these graphs, while giving the same information, the graphs better illustrates how much of the distribution our p-value takes up.




# Task 7

o	> t.test(x1,mu=23) # A 
o	
o	        One Sample t-test # B 
o	
o	data:  x1
o	t = 2.3563, df = 29, p-value = 0.02543 # C 
o	alternative hypothesis: true mean is not equal to 23 #D
o	95 percent confidence interval: #E
o	 23.30198 27.27320 #F
o	sample estimates:
o	mean of x 
o	 25.28759 #G

## A

Does the t test on sample x1 with a null hypothesis of mu equals 23.

## B

The name of the test.

## C

$$t_{calc}=\frac{\bar{x1}-\mu_0}{\frac{s}{\sqrt{n}}}$$
The different statistics that are calculated from the t.test. The t value which is calculated using the tcalc formula above. The degrees of freedom which is 30-1=29. The p-value which is the region below tcalc and above tcalc all subtracted by 1.

## D

The researchers hunch or alternative hypothesis where mu does not equal 23.

## E

Our confidence level with alpha=.05 since 1-.05=.95.

## F

Our 95% confidence interval.

## G

The estimated mean of x1.

# Task 8

```{r}
tcalc=(mean(x1)-25)/(sd(x1)/sqrt(30))
MATH4753Kelly::mypvalue(t0 = tcalc,n = 30,alpha = .05)
```

