---
title: "Lab1"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tasks

## Task 1

Find the working directory

```{r}
getwd()
```

## Task 2

Read in the DDT data.

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```

## Task 3

  a. The qualitative variables are RIVER and SPECIES
  b. The quantitative variables are MILE, DDT, LENGTH, and WEIGHT
  c. CCATFISH, SMBUFFALO, and LMBASS. There are three different species in the data set.
```{r}
with(ddt, table(SPECIES))
```
  
  d. 
```{r}
b800 <- with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS",])
b800
```
  e.
```{r}
scm4 <- with(ddt, ddt[RIVER=="SCM" & DDT>4.0,])
scm4
```
## Task 4

  a. Make a table of the rivers.
```{r}
riv=with(ddt, table(RIVER))
riv
```
  b. Make a color coded barplot of the rivers.
```{r}
barplot(riv, col = 1:4)
```
  
  c. Make a table of the rivers crossed with the species of fish.
```{r}
rivsp=with(ddt, table(RIVER,SPECIES))
rivsp
```
  d. Make a barplot of the rivers crossed with the species of fish.
```{r}
barplot(rivsp,beside = T, col = cm.colors(3))
```

## Task 5

  a. Make a pie chart of Fish SPECIES
```{r}
sp=with(ddt, table(SPECIES))
pie(sp, col = 5:7)
```
  
  b. Make a pie chart of RIVERS
```{r}
pie(riv, col=3:7)
```

## Task 6

  a. Make a boxplot of the variable DDT
  b. Make a boxplot of the variable WEIGHT
  c. Make a boxplot of the variable LENGTH
```{r}
layout(matrix(c(1,2,3), nrow = 1, ncol = 3))
with(ddt, boxplot(DDT, ylab = "DDT",col = "Green",notch=T))
with(ddt, boxplot(WEIGHT, ylab="WEIGHT",col="Yellow",notch=T))
with(ddt, boxplot(LENGTH, ylab="Length",col="Red",notch=T))
```
  
## Task 7

  a. Make a coplot of LENGTH Vs WEIGHT given RIVER
```{r}
fishcol = with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                          ifelse(SPECIES=="SMBUFFALO","Blue","Green")))
coplot(LENGTH~WEIGHT|RIVER, col = fishcol, data = ddt)
```
  
  b. Make a coplot of DDT Vs WEIGHT given SPECIES
```{r}
rivcol=with(ddt, ifelse(RIVER=="FCM","Red",
                        ifelse(RIVER=="LCM","Green",
                               ifelse(RIVER=="SCM","Blue","Black"))))
coplot(DDT~WEIGHT|SPECIES, col = rivcol, data=ddt)
```
  
## Task 8

  a. Use ggplot and make the following boxplots (be sure to put your name on the title using ggtitle())
```{r}
library(dplyr)
library(ggplot2)
ddt %>% ggplot(aes(x=SPECIES,y=WEIGHT,fill=RIVER)) + geom_boxplot() + ggtitle("Samuel Kelly")

```
  
  b. Use ggplot and make the following violin plots (name on the title)
```{r}
ddt %>% ggplot(aes(x=RIVER,y=LENGTH,fill=SPECIES)) + geom_violin() + ggtitle("Samuel Kelly")
```
  
  c. Use ggplot and make the following scatter plots (name on the title)
```{r}
ddt %>% ggplot(aes(x=WEIGHT,y=LENGTH,fill=SPECIES,color=SPECIES)) + geom_point() + ggtitle("Samuel Kelly")
```
  
## CLICKER QUESTIONS!

  1. The average length of the fish is 42.8125.
```{r}
mean(ddt$LENGTH)
```
  2. The standard deviation of the fish's weight is 376.5461.
```{r}
sd(ddt$WEIGHT)
```
  3. No, the length and width are on the opposite axis.
```{r}
plot(ddt$LENGTH, ddt$WEIGHT)
```
  
  
  4. It equals 1.00.
```{r}
v=1:20
v/20
```
  
  
$$R^2=\frac{MSS}{TSS}$$
