---
title: "Lab 10"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

```{r}
getwd()
```

# Task 2

```{r}
mymaxlik=function(lfun,x,param,...){
# how many param values are there?
np=length(param)
# outer -- notice the order, x then param
# this produces a matrix – try outer(1:4,5:10,function(x,y) paste(x,y,sep=" "))   to understand
z=outer(x,param,lfun) # A 
# z is a matrix where each x,param is replaced with the function evaluated at those values
y=apply(z,2,sum)

# y is a vector made up of the column sums
# Each y is the log lik for a new parameter value
plot(param,y,col="Blue",type="l",lwd=2,...)
# which gives the index for the value of y >= max.
# there could be a max between two values of the parameter, therefore 2 indices
# the first max will take the larger indice
i=max(which(y==max(y))) # B
abline(v=param[i],lwd=2,col="Red")

# plots a nice point where the max lik is
points(param[i],y[i],pch=19,cex=1.5,col="Black")
axis(3,param[i],round(param[i],2))
#check slopes. If it is a max the slope should change sign from + to 
# We should get three + and two -vs
ifelse(i-3>=1 & i+2<=np, slope<-(y[(i-2):(i+2)]-y[(i-3):(i+1)])/(param[(i-2):(i+2)]-param[(i-3):(i+1)]),slope<-"NA")
return(list(i=i,parami=param[i],yi=y[i],slope=slope))
}

```

## A

What outer does is that it takes the first argument and puts it in the first position of a matrix, then it takes the second argument and puts that in the first position  of the matrix as well. Then outer applies a function on those two values in the matrix, then continues with the rest of the values stored in x and param until we get a matrix of x rows and param columns.

## B

B stores in i the max index of y which y summed up all the columns.

## Formula

$$p(y_1,y_2,...,y_n)=p(y_1)...p(y_n)$$
$$p(y_1,y_2,...,y_8)=p(y_1)...p(y_8)$$
$$\text{The Binomial distribution formula is: }p(y)= {n \choose y}p^yq^{n-y}.$$
$$p(y_1,y_2,...,y_8)={20 \choose y_1}p^{y_1}(1-p)^{20-y_1}...{20 \choose y_8}p^{y_8}(1-p)^{20-y_8}$$

## R

```{r}
#dbinom(x = y1,size = 20,prob = p)*dbinom(x = y2,size = 20,prob = p)*dbinom(x = y3,size = 20,prob = p)*dbinom(x = y4,size = 20,prob = p)*dbinom(x = y5,size = 20,prob = p)*dbinom(x = y6,size = 20,prob = p)*dbinom(x = y7,size = 20,prob = p)*dbinom(x = y8,size = 20,prob = p)
```


## Graph

```{r}
logbin=function(x,param) log(dbinom(x,prob=param,size=20))
y = c(3,3,4,3,4,5,5,4)
mymaxlik(lfun = logbin,x = y,param=seq(0,1,length=1000), main = "Binomial, n = 20")
```

# Task 3

# Task 4

# Task 5

# Task 6

# Task 7

# Task 8
