---
title: "Lab 9"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

```{r}
getwd()
```

# Task 2

## Line A, Line B

```{r}
myboot<-function(iter=10000,x,fun="mean",alpha=0.05,...){  

#Notice where the ... is repeated in the code
n=length(x)   #sample size

#Now sample with replacement
y=sample(x,n*iter,replace=TRUE) #A

# Make a matrix with all the resampled values
rs.mat=matrix(y,nr=n,nc=iter,byrow=TRUE)
xstat=apply(rs.mat,2,fun) 
# xstat is a vector and will have iter values in it 
ci=quantile(xstat,c(alpha/2,1-alpha/2)) #B
# Nice way to form a confidence interval
# A histogram follows
# The object para will contain the parameters used to make the histogram
para=hist(xstat,freq=FALSE,las=1,main="Histogram of Bootstrap sample statistics",...)

#mat will be a matrix that contains the data, this is done so that I can use apply()
mat=matrix(x,nr=length(x),nc=1,byrow=TRUE)

#pte is the point estimate
#This uses whatever fun is
pte=apply(mat,2,fun)
abline(v=pte,lwd=3,col="Black")# Vertical line
segments(ci[1],0,ci[2],0,lwd=4)      #Make the segment for the ci
text(ci[1],0,paste("(",round(ci[1],2),sep=""),col="Red",cex=3)
text(ci[2],0,paste(round(ci[2],2),")",sep=""),col="Red",cex=3)

# plot the point estimate 1/2 way up the density
text(pte,max(para$density)/2,round(pte,2),cex=3)

return(list(ci=ci,fun=fun,x=x))# Some output to use if necessary
}
```

Line A assigns y the sample on a population of size n*iter, and we sample from the population with replacement that way we can get a distribution.

Line B assigns ci the quantiles of our y sample that had been turned into a matrix with some function applied to it. The quantile probabilities are alpha/2 and 1-alpha/2, so our theta lies somewhere in that interval.

## Why is replacement necessary?

To get a distribution we need to replace the values, since we will need to continually resample the data to due analysis on the data as well, like if there are errors in estimates of certain functions. The replacement keeps the size of the population we are sampling from consistent. 

## Unique replace

```{r}
set.seed(35) # This will give everyone the same sample
sam=round(rnorm(20,mean=10,sd=4),2)
unique(sample(sam,20,replace=TRUE) ) # repeat this line 5Xs
unique(sample(sam,20,replace=TRUE) )
unique(sample(sam,20,replace=TRUE) )
unique(sample(sam,20,replace=TRUE) )
unique(sample(sam,20,replace=TRUE) )
```

The function unique only lets us output unique values, so repeats are thrown out. This data that we have outputted tells us how many times a repeat was thrown out, since there should normally be an output of 20, but all of these unique functions are outputting a number less than 20.

## Unique w/o replace

```{r}
unique(sample(sam,20,replace=FALSE) )
unique(sample(sam,20,replace=FALSE) )
unique(sample(sam,20,replace=FALSE) )
unique(sample(sam,20,replace=FALSE) )
unique(sample(sam,20,replace=FALSE) )
```

Each one of these outputs are outputting the same exact sample, where all 20 numbers are unique, so no replacement occurs. The only real difference here is that the numbers are outputted in a different order in each case.

## 21

```{r}
#sample(sam,21,replace=FALSE) 
```

The sample size we are looking at is of size 20, and the above code is of size 21, and we cannot pull 21 things out of a hat with 20 items.

# Task 3

```{r}
myboot2<-function(iter=10000,x,fun="mean",alpha=0.05,cx=1.5,...){  #Notice where the ... is repeated in the code
n=length(x)   #sample size

y=sample(x,n*iter,replace=TRUE)
rs.mat=matrix(y,nr=n,nc=iter,byrow=TRUE)
xstat=apply(rs.mat,2,fun) # xstat is a vector and will have iter values in it 
ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
# A histogram follows
# The object para will contain the parameters used to make the histogram
para=hist(xstat,freq=FALSE,las=1,
main=paste("Histogram of Bootstrap sample statistics","\n","alpha=",alpha," iter=",iter,sep=""),
...)

#mat will be a matrix that contains the data, this is done so that I can use apply()
mat=matrix(x,nr=length(x),nc=1,byrow=TRUE)

#pte is the point estimate
#This uses whatever fun is
pte=apply(mat,2,fun)
abline(v=pte,lwd=3,col="Black")# Vertical line
segments(ci[1],0,ci[2],0,lwd=4)      #Make the segment for the ci
text(ci[1],0,paste("(",round(ci[1],2),sep=""),col="Red",cex=cx)
text(ci[2],0,paste(round(ci[2],2),")",sep=""),col="Red",cex=cx)

# plot the point estimate 1/2 way up the density
text(pte,max(para$density)/2,round(pte,2),cex=cx)
list(ci=ci,fun=fun,x=x)# Some output to use if necessary
}
```

## Plot mean

```{r}
set.seed(39)
sam = rnorm(25,mean = 25,sd = 10)
myboot(x = sam,alpha = .05,iter = 10000,xlab = "mean")
mean(sam)
```

```{r}
set.seed(30)
sam = rchisq(20,df = 3)
myboot(x = sam,iter = 10000,alpha = .05,xlab="mean")
mean(sam)
```

```{r}
set.seed(40)
sam = rgamma(30,shape = 2,scale = 3)
myboot(x = sam,iter = 10000,alpha = .05,xlab="mean")
mean(sam)
```

```{r}
set.seed(10)
sam = rbeta(20,shape1 = 3,shape2 = 4)
myboot(x = sam,iter = 10000,alpha = .05,xlab="mean")
mean(sam)
```

## Population mean

In each case we see that the population mean is very close to the mean that was estimated.
All of these estimates are in the confidence interval we are looking at.

## Plot variance

```{r}
set.seed(39)
sam=rnorm(25,mean = 25,sd = 10)
myboot2(x = sam,iter = 10000,alpha = .2,fun = var,xlab="variance")
```

```{r}
set.seed(30)
sam=rchisq(20,df = 3)
myboot2(x = sam,iter = 10000,alpha = .2,fun = var,xlab="variance")
```

```{r}
set.seed(40)
sam=rgamma(30,shape = 2,scale = 3)
myboot2(x = sam,iter = 10000,alpha = .2,fun = var,xlab="variance")
```

```{r}
set.seed(10)
sam=rbeta(20,shape1 = 3,shape2 = 4)
myboot2(x = sam,iter = 10000,alpha = .2,fun = var,xlab="variance")
```

# Task 4

# Task 5

## mean(x)/medium(x)

```{r}
mean.medium = function(x){
  mean(x)/median(x)
}
```


## 95% interval

```{r}
set.seed(39)
sam=rnorm(25,mean = 25,sd = 10)
myboot2(x = sam,iter = 10000,alpha = .05,fun = "mean.medium",xlab="mean(x)/medium(x)")
```

```{r}
set.seed(30)
sam=rchisq(20,df = 3)
myboot2(x = sam,iter = 10000,alpha = .05,fun = "mean.medium",xlab="mean(x)/medium(x)")
```

```{r}
set.seed(40)
sam=rgamma(30,shape = 2,scale = 3)
myboot2(x = sam,iter = 10000,alpha = .05,fun = "mean.medium",xlab="mean(x)/medium(x)")
```

```{r}
set.seed(10)
sam=rbeta(20,shape1 = 3,shape2 = 4)
myboot2(x = sam,iter = 10000,alpha = .05,fun = "mean.medium",xlab="mean(x)/medium(x)")
```

## 70% interval

```{r}
set.seed(39)
sam=rnorm(25,mean = 25,sd = 10)
myboot2(x = sam,iter = 10000,alpha = .3,fun = "mean.medium",xlab="mean(x)/medium(x)")
```

```{r}
set.seed(30)
sam=rchisq(20,df = 3)
myboot2(x = sam,iter = 10000,alpha = .3,fun = "mean.medium",xlab="mean(x)/medium(x)")
```

```{r}
set.seed(40)
sam=rgamma(30,shape = 2,scale = 3)
myboot2(x = sam,iter = 10000,alpha = .3,fun = "mean.medium",xlab="mean(x)/medium(x)")
```

```{r}
set.seed(10)
sam=rbeta(20,shape1 = 3,shape2 = 4)
myboot2(x = sam,iter = 10000,alpha = .3,fun = "mean.medium",xlab="mean(x)/medium(x)")
```

# Task 6

# Task 7

# Task 8

```{r}
MATH4753Kelly::myboot2(x = MATH4753Kelly::fd$DISTANCE,iter = 10000,fun = "mean",alpha = .05,xlab="mean")
```

