---
title: "Lab 14"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

```{r}
getwd()
```


# Task 2

## Function mylsq

```{r}
mylsq <- function(x,y){
  ssxx=sum((x-mean(x))*x )
  ssxy=sum((x-mean(x))*y) ## fill in the missing portion
  b1hat=ssxy/ssxx
  b0hat=mean(y)-b1hat*mean(x)   ##  fill in the missing portion
  return(list(b0hat=b0hat,b1hat=b1hat)) #fill in the missing portion
}
```

## Call the function

```{r}
set.seed(29)
x=1:20
y=4+6*x+rnorm(20,0,5)
lin=mylsq(x,y)
lin
```

## Plot

```{r}
plot(x = x,y = y,main = "Linear y vs x",xlab = "x: 1:20",ylab = "y: 4+6x+rnorm(20,0,5)")
abline(a = lin$b0hat,b = lin$b1hat, col="blue",lwd=2)
```

## SLR

```{r}
slr=lm(y~x)
summary(slr)
```

The manual calculation was $\hat\beta_0=3.102214$ and $\hat\beta_0= 6.130737$. The summary shows these exact calculations except with less accuracy.

# Task 3

## mypred()

```{r}
mypred=function(x,b0,b1){
ym=b0+b1*x ## fill in the gap
ym
}
```

## Predict 15.5

```{r}
mypred(15.5,lin$b0hat,lin$b1hat)
```

## Plot

```{r}
plot(x = x,y = y,main = "Linear y vs x",xlab = "x: 1:20",ylab = "y: 4+6x+rnorm(20,0,5)")
abline(a = lin$b0hat,b = lin$b1hat, col="blue",lwd=2)
points(cex=3,col="green",pch=19,x = 15.5,y=mypred(15.5,lin$b0hat,lin$b1hat))
```

## Data

```{r}
library(readxl)
oj=read_excel("OJUICE.xls")
dr=read_excel("DRILLROCK.xls")
head(oj)
head(dr)
```

## A OJ

```{r}
oj.lin=mylsq(oj$Pectin,oj$SweetIndex)
plot(x=oj$Pectin,y = oj$SweetIndex)
abline(a=oj.lin$b0hat,b = oj.lin$b1hat,col="blue",lwd=2)
```

## B OJ

```{r}
oj.lin
```

Here we can see that no pectin gives us an extremely sweet juice at 6.25 index, where the slope decreases as pectin is added, so when pectin is added sweetness decreases since the slope is -.0023.

## C OJ

```{r}
mypred(300,oj.lin$b0hat,oj.lin$b1hat)
```

## A DR

```{r}
dr.lin=mylsq(x = dr$DEPTH,y = dr$TIME)
plot(x=dr$DEPTH,y = dr$TIME)
```

## B DR

```{r}
mypred(x = dr$DEPTH,b0 = dr.lin$b0hat,b1 = dr.lin$b1hat)
```

## C DR

```{r}
plot(x=dr$DEPTH,y = dr$TIME)
abline(a=dr.lin$b0hat,b = dr.lin$b1hat,col="blue",lwd=2)
```

## D DR

```{r}
dr.lin
```

We can see that as depth increases so does time. We have a positive correlation between these variables. $\hat\beta_1$ has a slight positive slope, and $\hat\beta_0$ is 4.7896 at 0 depth.

# Task 4

## mysq function

```{r}
mysq=function(x,y){
n=length(x)
ssxx=sum((x-mean(x))*x )
ssxy=sum((x-mean(x))*y ) ## fill in the missing portion
b1hat=ssxy/ssxx
b0hat=mean(y)-b1hat*mean(x)   ##  fill 
yhat=b0hat+b1hat*x ##  fill 
ssr=sum((y-yhat)^2) # fill
sq= ssr/(n-2)##  fill 
return(list(ssr=ssr,sq=sq))
}
```

## Estimate sigma

```{r}
mysq(x = x,y = y)
anova(slr)
```

The values are very precise with residuals being 513.6~513.6199 and mean sq as 28.5~28.53444.

## A

```{r}
xv=c(1.6,2.4,2.8,2.9,3,3.1,3.2,3.2,3.3,3.3,3.4,3.4,3.8,4.3,3.6,3.4,3.3,3.2,3.2,3.4,3.3,3.2,3.1,3,2.7)
yv=c(1.7,3.1,3.3,3.6,3.4,3.5,3.8,3.7,3.5,3.4,3.6,3.5,3.9,3.4,3.2,3.3,3.1,3,2.9,2.6,2.5,2.6,2.6,2.4,2.6)
plot(x = xv,y = yv,ylab = "Pilot Plant",xlab = "Lab Furnace")
```

## B

```{r}
carb.lin=mylsq(xv,yv)
carb.lin
plot(x = xv,y = yv,ylab = "Pilot Plant",xlab = "Lab Furnace")
abline(a=carb.lin$b0hat,b = carb.lin$b1hat,col="blue",lwd=2)
```

We see that our slope is positive and our $\hat\beta_0$ is 1.265 when x is 0.

## C

```{r}
mysq(xv,yv)
carb.slr=lm(yv~xv)
anova(carb.slr)
```

The function correctly identified the residuals sum of squares and the mean squares.

## D

```{r}
sqrt(mysq(xv,yv)$sq)
```

This is the standard deviation between our means
