---
title: "Assignment 2"
author: "Samuel Kelly"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 12/17

# 1. [3 marks] MS 3.36 pg 105-106

## a. 

```{r}
noquote(paste0((1-.9212)*100,"% failure rate."))
```

## b.

```{r}
noquote(paste0((1-.7455)*100,"% failure rate."))
```

## c.

It's more likely that the novice will be the one who fails to identify the match as they have a 25.45% chance of failing while the experts have a 7.88% chance of failing.

# 2. [3 marks] MS 3.52 - pg 111

Population of 1000, 100 illegally using testosterone, 50 users test positive, 9 nonusers test positive.

## a.

```{r}
50/100
```

## b.

```{r}
(900-9)/900
```

## c.

P(USER|POSITIVE) = P(POSITIVE|USER)P(USER)/P(POSITIVE)
```{r}
noquote(paste0(round(((50/100)*(100/1000))/(59/1000)*100,4),"% chance that an athlete is really doping."))
```

# 3. [1 marks] MS Theorem 3.1 - pg 113 Prove the theorem in your own words.

## Theorem

You have k sets of elements, n1 for the first set, n2 for the second, nk for the kth set. Suppose you want to form a sample of k elements by taking one element from each of the k sets. The number of different samples is n1 * n2 ... nk.

## Proof



# 4. [1 marks] MS Theorem 3.2 - pg 114 Prove the theorem in your own words.

## Theorem

## Proof

# 5. [1 marks] MS Theorem 3.3 - pg 116 Prove the theorem in your own words.

## Theorem

## Proof

# 6. [1 marks] MS Theorem 3.4 - pg 117 Prove the theorem in your own words.

## Theorem

## Proof

# 7. [3 marks] MS 4.2 - pg 138

## a.

P(0)+P(1)+P(2)+P(3)+P(4)=1

```{r}
.09+.3+.37+.2+.04
```

## b.

P(3)+P(4)

```{r}
.2+.04
```

## c.

P(0)+P(1)

```{r}
.09+.3
```

# 8. [4 marks] MS 4.12 - pg 143

## Properties

$$0\leq P(y)\leq1\\\sum_{all ~y}P(y)=1$$

## a.

### Condition 1

All probabilities lie in the specified range.

### Condition 2

```{r}
.17+.1+.11+.11+.1+.1+.07+.05+.03+.02+.02+.02+.02+.02+.01+.01+.01+.01+.01+.005+.005
```

## b.

$$P(Y\geq 10)$$

```{r}
.02+.02+.02+.02+.01+.01+.01+.01+.01+.005+.005
```

## c.

### MEAN

```{r}
mu<-0*.17+1*.1+2*.11+3*.11+4*.1+5*.1+6*.07+7*.05+8*.03+9*.02+10*.02+11*.02+12*.02+13*.02+14*.01+15*.01+16*.01+17*.01+18*.01+19*.005+20*.005
mu
```

### VARIANCE

```{r}
variance <- (((0-mu)^2)*.17+((1-mu)^2)*.1+((2-mu)^2)*.11+((3-mu)^2)*.11+((4-mu)^2)*.1
+((5-mu)^2)*.1+((6-mu)^2)*.07+((7-mu)^2)*.05+((8-mu)^2)*.03+((9-mu)^2)*.02
+((10-mu)^2)*.02+((11-mu)^2)*.02+((12-mu)^2)*.02+((13-mu)^2)*.02+((14-mu)^2)*.01
+((15-mu)^2)*.01+((16-mu)^2)*.01+((17-mu)^2)*.01+((18-mu)^2)*.01+((19-mu)^2)*.005
+((20-mu)^2)*.005)
variance
```

## d.

```{r}
sigma <- sqrt(variance)
noquote(paste0("(",round((mu-2*sigma),4),",",round((mu+2*sigma),4),") will contain at least 75% of the probability."))
```

By this, if we take the interval (0,13) then we would see how much of the probability actually lies there.

```{r}
.17+.1+.11+.11+.1+.1+.07+.05+.03+.02+.02+.02+.02+.02
```

# 9. [4 marks] MS 4.34 - pg 154

n=25, foreign = .7, local = .3

## a.

```{r}
dbinom(10,25,.7)
```

## b.

```{r}
pbinom(5,25,.7)
```

## c.

### MEAN

n*p

```{r}
.7*25
```

### Standard Deviavtion

$$\sqrt {n * p * q}$$

```{r}
sqrt(25*.7*.3)
```

## d.

Out of our sample of 25 PhD graduates, roughly 17.5 of them are foreign students, and our deviation of 2.29 tells us that our data is close to the mean, so our standard deviation backs up our results.

# 10. [2 marks] MS 4.46 - pg 158

## a.

```{r}
dmultinom(x = c(5,5,5,5,5,5,5,5,5,5),prob = c(.1,.1,.1,.1,.1,.1,.1,.1,.1,.1))
```

## b.

```{r}
pbinom(1,50,.1)
```

# 11. [2 marks] MS 4.54 - pg 162

## a.

$${y-1 \choose 1-1}*.18^1*.82^1$$

## b.

```{r}
noquote(paste0("Roughly ",round(1/.18,4), " or 6 consumers until one gives an answer of other."))
```

## c.

```{r}
dnbinom(x = 1-1,size = 1,prob = .18)
```

## d.

```{r}
1-pnbinom(q = 1,size = 1,prob = .18)
```

# 12. [2 marks] MS 4.66 - pg 168

## a.

N=209,n=10,r=8

```{r}
10*(8/209)
```

So .3828 is the amount of facilities we expect to be able to treat hazardous waste on-site. So in a random sample of 10 out of 209, we have close to 0 chance of even getting 1 in our sample.

## b.

```{r}
dhyper(x = 4,m = 8,n = 201,k = 10)
```

# 13. [3 marks] MS 4.78 - pg 173

## a.

$$\mu=\sigma^2=.03$$

## b.

We assume that the events are independent of each other, which they are since one boat sinking and another boat sinking at different place and time within 3 years have no correlation to each other. Also since these are rare events we use a Poisson distribution which allows us to use the mean and variance as the same value.

## c.

```{r}
noquote(paste0(round(dpois(0,.03),4)," or 97.04% chance of 0 casualties in a 3 year period."))
```


# 14. [4 marks] MS 5.2 - pg 191

$$f(y)=c(2-y)  \text{ if 0 } \leq{y} \leq 1 \\ 0  ~~~~~\text{ elsewhere}$$

## a.

Since the integration of f(y) spans -infinity to infinity, but only 0 to 1 has an expression worth integrating.

$$\int_{0}^{1}c(2-y)dy=1$$
$$c\int_{0}^{1}(2-y)dy=1$$
$$c(\int_{0}^{1}2dy-\int_{0}^{1}ydy)=1$$
$$c(2y|_{0}^{1}-\frac{y^2}{2}|_{0}^{1})=1$$
$$c(2(1)-\frac{1^2}{2})=1$$
$$c(\frac{3}{2})=1$$
$$c=\frac{3}{2}$$

$$f(y)=\frac{3}{2}(2-y)  \text{ if 0 } \leq{y} \leq 1 \\ 0  ~~~~~\text{ elsewhere}$$

## b.

$$F(y)=\int_{-\infty}^yf(t)dt$$
$$F(y)=\int_{0}^y\frac{3}{2}(2-t)dt$$
$$F(y)=\frac{3}{2}\int_{0}^y(2-t)dt$$
$$F(y)=(3t-\frac{3t^2}{4})|_0^y$$
$$F(y)=(3y-\frac{3y^2}{4})$$
$$F(y)=0 ~~~\text{ -}\infty < y <0,\\(3y-\frac{3y^2}{4})~~~~~0\leq y\leq1, \\0~~~~1<y$$

## c.

F(.4)

```{r}
(3*.4)-((3*.4^2)/4)
```

## d.

F(.6) - F(.1)

```{r}
(3*.6)-((3*.6^2)/4) - (3*.1)-((3*.1^2)/4)
```


# 15. [3 marks] MS 5.10 - pg 196

## a.

$$\mu=E(Y)=\int_{-5}^5yf(y)dy=\frac{3}{500}\int_{-5}^5(25y-y^3)dy=\frac{3}{500}(\frac{25y^2}2-\frac{y^4}{4})|_{-5}^5$$

```{r}
noquote(paste0("Mean is ",(3/500)*((((25*5^2)/2)-(5^4)/4)-(((25*(-5)^2)/2)-((-5)^4)/4)),"."))
```

$$E(Y^2)=\int_{-5}^5y^2f(y)dy=\frac{3}{500}\int_{-5}^5(25y^2-y^4)dy=\frac{3}{500}(\frac{25y^3}3-\frac{y^5}{5})|_{-5}^5$$

```{r}
(3/500)*((((25*5^3)/3)-(5^5)/5)-(((25*(-5)^3)/3)-((-5)^5)/5))
```

$$\sigma^2=E(Y^2)-\mu^2=5-0^2=5$$

## b.

### MEAN

```{r}
0*(1/60)
```

### VARIANCE

```{r}
5*(1/3600)
```

## c.

### MEAN

```{r}
0*60
```

### VARIANCE 

```{r}
3600*5
```


# 16. [3 marks] MS 5.36 - pg 205

mean 50, sd 3.2

## a.

```{r}
1-pnorm(45,50,3.2)
```

## b.

```{r}
pnorm(55,50,3.2)
```

## c.

```{r}
pnorm(52,50,3.2)-pnorm(51,50,3.2)
```


# 17. [5 marks] MS 5.38- pg 205

35 mph

```{r}
crash <- read.csv("CRASH.csv")
```

