---
title: "Assignment 2"
author: "Samuel Kelly"
date: '`r format(Sys.Date(),format="%A, %B %d, %Y")`'
output: 
  html_document:
    df_print: paged
    fig_caption: true
    highlights: pygments
    number_sections: no
    theme: spacelab
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. [3 marks] MS 3.36 pg 105-106

## a. 

```{r}
noquote(paste0((1-.9212)*100,"% failure rate."))
```

## b.

```{r}
noquote(paste0((1-.7455)*100,"% failure rate."))
```

## c.

It's more likely that the novice will be the one who fails to identify the match as they have a 25.45% chance of failing while the experts have a 7.88% chance of failing.

# 2. [3 marks] MS 3.52 - pg 111

Population of 1000, 100 illegally using testosterone, 50 users test positive, 9 nonusers test positive.

## a.

```{r}
50/100
```

## b.

```{r}
(900-9)/900
```

## c.

P(USER|POSITIVE) = P(POSITIVE|USER)P(USER)/P(POSITIVE)
```{r}
noquote(paste0(round(((50/100)*(100/1000))/(59/1000)*100,4),"% chance that an athlete is really doping."))
```

# 3. [1 marks] MS Theorem 3.1 - pg 113 Prove the theorem in your own words.

## Theorem

You have k sets of elements, n1 for the first set, n2 for the second, nk for the kth set. Suppose you want to form a sample of k elements by taking one element from each of the k sets. The number of different samples is n1 * n2 ... nk.

## Proof



# 4. [1 marks] MS Theorem 3.2 - pg 114 Prove the theorem in your own words.

## Theorem

## Proof

# 5. [1 marks] MS Theorem 3.3 - pg 116 Prove the theorem in your own words.

## Theorem

## Proof

# 6. [1 marks] MS Theorem 3.4 - pg 117 Prove the theorem in your own words.

## Theorem

## Proof

# 7. [3 marks] MS 4.2 - pg 138

$$2\frac{\mu g}{g}$$
Random sample of 4 homes

## a.

P(0)+P(1)+P(2)+P(3)+P(4)=1

```{r}
.09+.3+.37+.2+.04
```

## b.

P(3)+P(4)

```{r}
.2+.04
```

## c.

P(0)+P(1)

```{r}
.09+.3
```


# 8. [4 marks] MS 4.12 - pg 143

# 9. [4 marks] MS 4.34 - pg 154

# 10. [2 marks] MS 4.46 - pg 158

# 11. [2 marks] MS 4.54 - pg 162

# 12. [2 marks] MS 4.66 - pg 168

# 13. [3 marks] MS 4.78 - pg 173

# 14. [4 marks] MS 5.2 - pg 191

# 15. [3 marks] MS 5.10 - pg 196

# 16. [3 marks] MS 5.36 - pg 205

# 17. [5 marks] MS 5.38- pg 205
