---
title: "Lab 7"
author: "Samuel Kelly"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

## Working Directory

```{r}
getwd()
```

# Task 2

## mychisim Function

```{r}
mychisim<-function(n1=10,sigma1=3,mean1=5,iter=1000,ymax=0.1,x=20, y=0.1){    # adjust ymax to make graph fit
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
#n1*iter is the size of the matrix where each iter column is a new sample of the population
data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1

ssq1=apply(data1.mat,2,var) # ssq1 is s squared

w=(n1-1)*ssq1/sigma1^2      #chi-sq stat

hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]," = ",n1," statistic = ",chi^2)),
xlab=expression(paste(chi^2, "Statistic",sep=" ")), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(dchisq(x,n1-1),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(chi^2==frac((n[1]-1)*s^2,sigma^2)) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
invisible(list(w=w,summary=summary(w),sd=sd(w),fun="Chi-sq")) # some output to use if needed
}
```

## 4 Chisq plots

```{r}
mychisim(n1 = 10,iter = 1000,mean1 = 10,sigma1 = 4,x = 20,y = .06,ymax = .11)
mychisim(n1 = 20,iter = 1000,mean1 = 10,sigma1 = 4,x = 30,y = .04,ymax = .07)
mychisim(n1 = 100,iter = 1000,mean1 = 10,sigma1 = 4,ymax = .03,x = 130,y = .020)
mychisim(n1 = 200,iter = 1000,mean1 = 10,sigma1 = 4,ymax = .02,x = 240,y = .015)
```

## Histogram of w

```{r}
chisq = mychisim(n1 = 10,iter = 1500,mean1 = 20,sigma1 = 10,ymax = .11,x = 20,y = .08)
hist(chisq$w,main = "Histogram of Chisq",col = rainbow(15),xlab = expression(paste(chi^2," Value")))
```

# Task 3

## myTsim function

```{r}
myTsim<-function(n1=10,sigma1=3,mean1=5,iter=1000,ymax=0.1,x=2,y=0.3,...){    # adjust ymax to make graph fit
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1

data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1

sd1=apply(data1.mat,2,sd) # sd
ybar=apply(data1.mat,2,mean)  # mean

w=(ybar-mean1)/(sd1/sqrt(n1))      #T stat

hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]," = ",n1," statistic = ",T," iterations= ",iter)),
xlab=expression(paste(T, "Statistic",sep=" ")), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(dt(x,n1-1),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(T==frac((bar(y)-mu),s/sqrt(n1))) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
invisible(list(w=w,summary=summary(w),sd=sd(w),fun="T")) # some output to use if needed
}
```

## 4 plots

```{r}
myTsim(n1 = 10,iter = 1000,mean1 = 10,sigma1 = 4,ymax = .4)
myTsim(n1 = 20,iter = 1000,mean1 = 10,sigma1 = 4,ymax = .4)
myTsim(n1 = 100,iter = 1000,mean1 = 10,sigma1 = 4,ymax = .4,x = -3.5)
myTsim(n1 = 200,iter = 1000,mean1 = 10,sigma1 = 4,ymax = .4,x = 1.5)
```

## Histogram of w

```{r}
T <- myTsim(n1 = 10,iter = 1500,mean1 = 20,sigma1 = 10,ymax = .4)
hist(T$w,xlab = "T Value",col = rainbow(15))
```

# Task 4

## 2 population mychisim function

```{r}
mychisim2<-function(n1=10,n2=14,sigma1=3,sigma2=3,mean1=5,mean2=10,iter=1000,ymax=0.07,x=40,y=0.04,...){    # adjust ymax to make graph fit
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
ssq1=apply(data1.mat,2,var) # ssq1 is s squared
ssq2=apply(data2.mat,2,var)
spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
w=(n1+n2-2)*spsq/(sigma1^2)#sigma1=sigma2,  Chi square stat
hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",chi^2)),
xlab=expression(paste(chi^2, "Statistic",sep=" ")), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(dchisq(x,n1+n2-2),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(chi^2==frac((n[1]+n[2]-2)*S[p]^2,sigma^2)) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title) # Legend #
invisible(list(w=w,summary=summary(w),sd=sd(w),fun="Chi-sq")) # some output to use if needed
}
```

## Parameterized output

```{r}
mychisim2(n1 = 10,n2 = 10,mean1 = 5,mean2 = 10,sigma1 = 4,sigma2 = 4,iter = 1000,ymax = .08,x = 30)
mychisim2(n1 = 20,n2 = 10,mean1 = 3,mean2 = 5,sigma1 = 10,sigma2 = 10,iter = 1000,ymax = .06,x = 40)
mychisim2(n1 = 50,n2 = 50,mean1 = 5,mean2 = 10,sigma1 = 4,sigma2 = 4,iter = 10000,ymax = .03,x = 120,y = .02)
mychisim2(n1 = 80,n2 = 50,mean1 = 3,mean2 = 5,sigma1 = 10,sigma2 = 10,iter = 10000,ymax = .03,x = 160,y = .015)
```

## Default output Histogram with iter=10000

```{r}
chisq2.pop <- mychisim2(iter = 10000)
hist(chisq2.pop$w,col = rainbow(15),xlab = expression(paste(chi^2," Value")))
```

# Task 5

## 2 population myTsim function

```{r}
myTsim2<-function(n1=10,n2=14,sigma1=3,sigma2=3,mean1=5,mean2=10,iter=1000,ymax=0.5,x=2,y=0.4,...){
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
ssq1=apply(data1.mat,2,var) # ssq1 is s squared
ybar1= apply(data1.mat,2,mean)
ssq2=apply(data2.mat,2,var)
ybar2=apply(data2.mat,2,mean)
spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
w=((ybar1-ybar2)-(mean1-mean2))/sqrt(spsq*(1/n1+1/n2))#sigma1=sigma2,  Chi square stat
hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",T)),
xlab=paste(" T Statistic",sep=""), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(dt(x,n1+n2-2),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(T==frac((bar(Y)[1]-bar(Y)[2])-(mu[1]-mu[2]),S[p]*sqrt(frac(1,n[1])+frac(1,n[2])))) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)# Legend #
invisible(list(w=w,summary=summary(w),sdw=sd(w),fun="T")) # some output to use if needed
}
```

## T Statistic

$$t=\frac{\bar{y}-\mu}{S/\sqrt{n}}$$
$$\bar{y}~is~the~normally~distributed~random~variable$$
$$\mu ~is~the~expected~value~of~\bar{y}$$
$$S~is~the~variance$$
$$n ~is~ the~ population$$

## 4 plots

```{r}
myTsim2(n1 = 10,n2 = 10,mean1 = 5,mean2 = 10,sigma1 = 4,sigma2 = 4,iter = 1000,ymax = .45,x = 1.5)
myTsim2(n1 = 20,n2 = 10,mean1 = 3,mean2 = 5,sigma1 = 10,sigma2 = 10,iter = 1000,ymax = .45,x = 1)
myTsim2(n1 = 50,n2 = 50,mean1 = 5,mean2 = 10,sigma1 = 4,sigma2 = 4,iter = 10000,ymax = .45,x = 1)
myTsim2(n1 = 80,n2 = 50,mean1 = 3,mean2 = 5,sigma1 = 10,sigma2 = 10,iter = 10000,ymax = .45,x = 1)
```

## Histogram of w

```{r}
T <- myTsim2(iter = 10000)
hist(T$w,col = rainbow(15),xlab = "T Values")
```


# Task 6

## 2 population myFsim function

```{r}
myFsim2<-function(n1=10,n2=14,sigma1=3,sigma2=2,mean1=5,mean2=10,iter=1000,ymax=0.9,x=6,y=0.5,...){
y1=rnorm(n1*iter,mean=mean1,sd=sigma1)# generate iter samples of size n1
y2=rnorm(n2*iter,mean=mean2,sd=sigma2)
data1.mat=matrix(y1,nrow=n1,ncol=iter,byrow=TRUE) # Each column is a sample size n1
data2.mat=matrix(y2,nrow=n2,ncol=iter,byrow=TRUE)
ssq1=apply(data1.mat,2,var) # ssq1 is s squared
ssq2=apply(data2.mat,2,var)
#spsq=((n1-1)*ssq1 + (n2-1)*ssq2)/(n1+n2-2) # pooled s squared
w=ssq1*sigma2^2/(ssq2*sigma1^2) #
hist(w,freq=FALSE, ylim=c(0,ymax), # Histogram with annotation
main=substitute(paste("Sample size = ",n[1]+n[2]," = ",n1+n2," statistic = ",F)),
xlab=paste("F Statistic",sep=""), las=1)
lines(density(w),col="Blue",lwd=3) # add a density plot
curve(df(x,n1-1,n2-1),xlim=c(0,6),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
title=expression(F==frac(s[1]^2,s[2]^2)*frac(sigma[2]^2,sigma[1]^2)) #mathematical annotation -see ?plotmath
legend(x,y,c("Simulated","Theoretical"),col=c("Blue","Red"),lwd=4,lty=1:2,bty="n",title=title)# Legend #
invisible(list(w=w,summary=summary(w),sd=sd(w),fun="F")) # some output to use if needed
}
```

## F statistics

The S is the sample variance, and the sigma is the population variance. Since this is a product of ratios, we do not need the variances to equal each other, otherwise we will simply get a 1 for one or both terms.

## Parameter

```{r}
myFsim2(n1 = 10,n2 = 10,mean1 = 5,mean2 = 10,sigma1 = 5,sigma2 = 10,iter = 1000,x = 4)
myFsim2(n1 = 20,n2 = 10,mean1 = 10,mean2 = 5,sigma1 = 2,sigma2 = 20,iter = 1000,x = 4)
myFsim2(n1 = 50,n2 = 50,mean1 = 5,mean2 = 10,sigma1 = 1,sigma2 = 2,iter = 10000,x = 2.5,ymax = 1.5)
myFsim2(n1 = 80,n2 = 50,mean1 = 3,mean2 = 5,sigma1 = 15,sigma2 = 5,iter = 10000,x = 2.5,ymax = 1.7)
```

## Histogram of w

```{r}
F <- myFsim2(iter = 10000)
hist(F$w,col = rainbow(15),xlab = "F Values")
```

# Task 7

```{r}
library(MATH4753Kelly)
data(fd)
knitr::kable(head(fd))
```
## ERROR Encountered

Error: 'fire' is not an exported object from 'namespace:MATH4753Kelly'

## The fire.R code in the R script file
```{r}
#' @title Fire data set.
#'
#' @description  A data set concerning damage from fire
#'
#' @details This is a standard data set from the course MATH 4753 taken from the data sets provided by the text book.
#'
#' @format A data frame with 15 rows and 2 variables:
#' \describe{
#'   \item{DISTANCE}{Distance in miles from fire station}
#'   \item{DAMAGE}{Thouands of dollars}
#' }
#'
#' @source \url{https://www.routledge.com/Statistics-for-Engineering-and-the-Sciences-Sixth-Edition/Mendenhall-Sincich/p/book/9781498728850}
"fire"
```

## NAMESPACE
When I remove fire.R from the R folder, the package builds just fine, but when I add this code above into that R folder in fire.R, it gives me the error above. I used  this code to get the right output, but I cannot use any functions that I have in my package, just the data that I just forced in. devtools::install(build_vignettes= TRUE, force = TRUE)
